### 2021-01-06 ch/2-related-work.tex
Finish first draft of chapter 3 (now 4)

\chapter{Related work}\label{ch:related-work}
%
% =========================
%      old introduction
% =========================
% \section{Introduction}\label{introduction}
% Fairness in Machine Learning is a topic that has recently received increasing attention.
% Machine Learning has become very widespread in our society
% which brings some dangers with it.
% Decisions such as credit eligibility and hiring are increasingly done automatically.
% However, the algorithms that are used in these decisions can have strong biases
% that stem from the training data.
% Often, the algorithms are even more biased than the training data.
% \citet{kamishima2011fairness} give a great illustration for this
% (adapted from \citet{calders2010three}).
% The illustration is based on the Adult/Census Income dataset \citep{kohavi1996scaling}
% which contains records of individuals with thirteen attributes, among them age and gender.
% The task is to predict whether or not the individual earns more than \$50K per year.
% In the dataset, 32\% of the men have a positive label (\ie, they earn more than \$50K),
% but of the women only 11\% have a positive label.
% With this ratio of about 3:1, the dataset is already skewed.
% However, when training a \href{https://en.wikipedia.org/wiki/Naive_Bayes_classifier}{Na\"ive Bayes classifier} on the data,
% the ratio in the predictions is even more unfair:
% the amount of women who get a positive prediction is even lower than 11\%.
% If the Na\"ive Bayes classifier does not get access to the gender information,
% the predictions will be a bit fairer,
% but they are still more unfair than the original dataset labels.
%
% This phenomenon of predictions being more unfair than an already skewed dataset
% is due to the generalization that classifiers have to make.
% It is easier to classify all but the most outstanding women as low income
% than to learn a complicated decision boundary.
% Furthermore, removing the gender information does not solve the problem
% because the gender can be inferred from other data.
%
% The following is a review of the literature on Fairness in Machine Learning.
% Several different approaches to and definition of Fairness are presented.
% The focus is on identifying the major trends in the field from the beginning.
\section{Two views of the dataset bias problem}\label{two-views-of-the-dataset-bias-problem}
As hinted to in the introduction,
there is a divide in the relevant literature in how the problem of dataset bias is approached.
On the one hand, there is the ``groundtruth-centric'' view -- expressed through most of the introduction --
that the training set is biased but the deployment setting is not.
By \emph{deployment setting} I am here referring to the real distribution
that will be encountered by the machine learning system once it is in use.

>>>>

\chapter{Related work}\label{ch:related-work}
%
% =========================
%      old introduction
% =========================
% \section{Introduction}\label{introduction}
% Fairness in Machine Learning is a topic that has recently received increasing attention.
% Machine Learning has become very widespread in our society
% which brings some dangers with it.
% Decisions such as credit eligibility and hiring are increasingly done automatically.
% However, the algorithms that are used in these decisions can have strong biases
% that stem from the training data.
% Often, the algorithms are even more biased than the training data.
% \citet{kamishima2011fairness} give a great illustration for this
% (adapted from \citet{calders2010three}).
% The illustration is based on the Adult/Census Income dataset \citep{kohavi1996scaling}
% which contains records of individuals with thirteen attributes, among them age and gender.
% The task is to predict whether or not the individual earns more than \$50K per year.
% In the dataset, 32\% of the men have a positive label (\ie, they earn more than \$50K),
% but of the women only 11\% have a positive label.
% With this ratio of about 3:1, the dataset is already skewed.
% However, when training a \href{https://en.wikipedia.org/wiki/Naive_Bayes_classifier}{Na\"ive Bayes classifier} on the data,
% the ratio in the predictions is even more unfair:
% the amount of women who get a positive prediction is even lower than 11\%.
% If the Na\"ive Bayes classifier does not get access to the gender information,
% the predictions will be a bit fairer,
% but they are still more unfair than the original dataset labels.
%
% This phenomenon of predictions being more unfair than an already skewed dataset
% is due to the generalization that classifiers have to make.
% It is easier to classify all but the most outstanding women as low income
% than to learn a complicated decision boundary.
% Furthermore, removing the gender information does not solve the problem
% because the gender can be inferred from other data.
%
% The following is a review of the literature on Fairness in Machine Learning.
% Several different approaches to and definition of Fairness are presented.
% The focus is on identifying the major trends in the field from the beginning.
\section{\NoCaseChange{Two views of the dataset bias problem}}\label{two-views-of-the-dataset-bias-problem}
As hinted to in the introduction,
there is a divide in the relevant literature in how the problem of dataset bias is approached.
On the one hand, there is the ``groundtruth-centric'' view -- expressed through most of the introduction --
that the training set is biased but the deployment setting is not.
By \emph{deployment setting} I am here referring to the real distribution
that will be encountered by the machine learning system once it is in use.

### 2021-01-06 ch/3-target-labels.tex
Finish first draft of chapter 3 (now 4)

\chapter{Tuning Fairness by Balancing Target Labels}\label{ch:target-labels}

>>>>

\chapter{Tuning Fairness by Balancing Target Labels}\label{ch:target-labels}
This chapter is predominantly concerned with label bias.
The main idea is that we make use of pseudo labels to implicitly learn from a \emph{balanced} dataset,
in which \(y\perp s\) holds, and which thus satisfies \acf{DP}.
This falls under the area of \emph{fair classifiers} discussed in \secref{fair-classifiers}

We can interpret the contributions of this chapter in two ways.
The first corresponds to the definition-centric view of dataset bias,
and the second to the groundtruth-centric view.
\begin{enumerate}
  \item
    We can say that the classifier should satisfy demographic parity in its predictions,
    and learning from a balanced training set is just one particular way to achieve this.
    In this view, the pseudo labels have no deeper meaning and are just a computational trick.
  \item
    We can see the training set as a corrupted version of a true dataset, which is balanced (\(y\perp s\)),
    and so, by learning from these pseudo labels, we are simply approximating the true dataset.
    However, we do not actually have access to the true dataset; we only know that it is balanced.
    In order to evaluate the trained model, we compute fairness metrics with respect to \ac{DP}.
\end{enumerate}
Within the chapter, we sometimes jump between these two views.

While the main focus is on label bias,
the experiments are performed on real-world fairness datasets,
which also display a significant amount of sample bias (disadvantaged groups are underrepresented).
Furthermore, in addition to the result for \ac{DP}, we also show that the proposed scheme improves \acf{EOpp}.

In order to construct the pseudo labels,
we use side information about summary statistics for an unbalanced training set.
This allows us to target a specific balanced set, instead of just any balanced set.
In other words, rather than just enforcing \ac{DP}, the method gives control over the target rates \(P(\hat{y}=1|s)\).
